---
title: "hw1"
author: "Karthik Ranganathan"
date: "January 30, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

6) Install the maps package (you do not need to load it) and use the ggplo2::map_data function to get data
## for drawing the "Four Corners" region of the United States (i.e., Arizona, New Mexico, Utah, and Colorado).
## Install the measurements package and use the measurements::conv_unit function to convert the latitude
## and longitude information in the dataset to decimal degrees suitable to be used for plotting.
## Plot a map of the region (you may want to adjust the plotting limits to an appropriate "zoom" level), and
## overlay the locations of the water sampling sites on the map. Use color to indicate the EPA Section and size
## to indicate the amount of Uranium-238 measured at each site.
```{r}

install.packages("maps")
library(maps)
install.packages("measurements")
library(measurements)
install.packages("mapproj")
library(mapproj)

map_four_state <- map_data("state",
                         region=c("colorado","new mexico",
                                  "utah","arizona")) 



novajo$Latitude = as.numeric(measurements::conv_unit(novajo$Latitude,
                                                "deg_dec_min",
                                                "dec_deg"))
novajo$Latitude

novajo$Longitude = as.numeric(measurements::conv_unit(novajo$Longitude,
                                                 "deg_dec_min",
                                                 "dec_deg"))
novajo$Longitude
novajo$Longitude = novajo$Longitude * (-1)
```

```{r}

ggplot(data = novajo) +
  geom_point(mapping = aes( x= Longitude, y = Latitude, color = Which.EPA.Section.is.This.From., size = Amount.of.Uranium238)) +
  geom_polygon(data = map_four_state, mapping = aes(x = long,
                                         y = lat,
                                         group = group),
               fill= NA,
               color="black")
```

